---
title: "ES 207 Homework 7"
output: html_document
---

Part I

 For the Water Years 2005-2012, develop a predictive model of the mean monthly Chl-a concentration in the California Bay Delta using  other mean monthly water quality variables.

Be sure to show your work, including multiple model comparison (or step models), provide quantification and explain the reasoning behind your final model selection.
Which variables are highly correlated with the others? How did you account for this?
Be sure to demonstrate that you have accounted for all assumptions in your regression. Identify any badly fitted observations, discuss whether your residuals are following expected patterns, and discuss whether there are any possible explanations for any poorly modeled observations.
Quantify how much variance in mean monthly Chl-a is explained by the variables you selected. 
```{r}
require(tidyverse)
wq <- read_csv("BayDeltaWQ.csv", col_names = TRUE, na = c("NA", "n/p"), guess_max = 30000)
```

```{r}

par(mfrow=c(2,4))

#First I play around with my data set a bit to see if I notice any correlations between chlorophyll a and other variables based I what I know about chlorophyll
plot(wq$`Chlorophyll a`)
plot(wq$`1% Light Depth`, wq$`Chlorophyll a`)
plot(wq$Depth, wq$`Chlorophyll a`)
#Since chlorophyll is a light capturing component of plant cells I would expect less of it to exist at greater depths where sunlight penentration decreases  It's usuless for plant cells to have more chlorophyll in low light conditions waste of resources and energy
plot(wq$`Manganese (Dissolved)`,wq$`Chlorophyll a`)
plot(wq$`Manganese (Total)`, wq$`Chlorophyll a`)
#Magnesium is a major building block for chlorophyll in general, so the more of that that is readily avaialable for plant cells to take in and mineralize, the more resources they have to create more chlorphyll if necessary
plot(wq$`Ammonia (Total)`, wq$`Chlorophyll a`) #High pH and ammonia has a negative effect on photosynthesis and can damage plant cells.  Whule plants can take in ammonia as a form of nitrogen, it needs to be materialized rapidly as ammonia in its current form is toxic to the plant cell.  I feel in an environment where pH is high and ammonia concentrations are high, than plant cells will not thrive as well and major die offs are a possibility leading to a decrease in Chlrophyll
plot(wq$pH, wq$`Chlorophyll a`)
plot(wq$Oxygen, wq$`Chlorophyll a`)
#Since within the chlorophyll during light dependent reactions, water gets split into hydrogen and O2 so I feel the greater concentration of plant cells and chlorophyll the more oxygen that should be released

#While some of these plots are showing up blank for me during my pre-tests, I definitely can notice signs of relationships and correlations between some of the variables like depth and magnesium vs. chlorophyll concentrations.  These are variables I will look into using for my final predictive model
```



```{r}
#Next I need to clean up my data by removing rows that include the sample dates that don't fit into the 2005-2012 range 

```

install.packages("dplyr")
```{r}
require(dbplyr)
wqs<-wq[12645:14958,]
wqt<-wq[28122:30158,]

#Monthly data for 2005-2012

```

```{r}
wq1<-wq[1:12644,]
wq2<-wq[14959:28121,]


#Monthly data for all other years outside of 2005-2012

#I also realize there are two separate sets of date ranges which I indicated a wq1 or wq2, one has certain data types that do not exist in the other one

#wq2 appears to have data in the variables I'm looking at such as Depth, Manganese, and Chlorophyll a

```

```{r}
require(tidyverse)
require(dplyr)

wq2com<-wq2[c(2,4,26,45,69)]

#Subsetting the variables I care about dependent on how correlated my plots seem as well as confrimation with R-squared value in relation to my knowledge about plant cells and chlorophyll

separate(wq2com, "SampleDate", c("Month", "Day", "Year"), sep = "/")

#I wanted to separate my SampleDate into Month, Day, and Year, to make it easier for sorting later


```

```{r}
CD.lm<-lm(wq$`Chlorophyll a`~wq$Depth)
summary(CD.lm)

CM.lm<-lm(wq$`Chlorophyll a`~wq$`Manganese (Dissolved)`)
summary(CM.lm)

CA.lm<-lm(wq$`Chlorophyll a`~wq$`Ammonia (Total)`)
summary(CA.lm)

#Running linear models of my main predicted variables in realtion to chlorophyll a concentrations (water depth, manganese, and ammonia)

```


```{r}
par(mfrow=c(1,3))

plot(wq2com$Depth, wq2com$`Chlorophyll a`)

plot(wq2com$`Ammonia (Total)`, wq2com$`Chlorophyll a`)

plot(wq2com$`Manganese (Dissolved)`, wq2com$`Chlorophyll a`)

#Replotting all my main variables for cleaner comparison


```

```{r}

ggplot(wq2com, aes(x = log(wq2com$Depth), y = wq2com$`Chlorophyll a`)) + geom_point()

ggplot(wq2com, aes(x = log(wq2com$`Manganese (Dissolved)`), y = wq2com$`Chlorophyll a`)) + geom_point()

ggplot(wq2com, aes(x = log(wq2com$`Ammonia (Total)`), y = wq2com$`Chlorophyll a`)) + geom_point()

#Data clean uo and log transform to normalize data points

```

```{r}
#Newer regression models using the cleaner data

CvD <- lm(log(wq2com$`Depth`)~wq2com$`Chlorophyll a`, data = wq2com)
summary(CvD)
CvA <- lm(log(wq2com$`Ammonia (Total)`)~wq2com$`Chlorophyll a`, data = wq2com)
summary(CvA)
CvM <- lm(log(wq2com$`Manganese (Dissolved)`)~wq2com$`Chlorophyll a`, data = wq2com)
summary(CvM)

```

install.packages("lubridate")
install.packages("modelr")
```{r}

require(modelr)
?modelr::add_predictions

wq2com <- wq2com %>% 
  spread_predictions(CvD, CvM, CvA)
wq2com

#Compare predictions

```

```{r}
ggplot(wq2com, aes(x = wq2com$`Ammonia (Total)`, y = exp(CvA))) + geom_point()+stat_smooth(method = "lm", se = F) + geom_abline(linetype = "dashed")

ggplot(wq2com, aes(x = wq2com$`Manganese (Dissolved)`, y = exp(CvM))) + geom_point()+stat_smooth(method = "lm", se = F) + geom_abline(linetype = "dashed")

ggplot(wq2com, aes(x = wq2com$Depth, y = exp(CvD))) + geom_point()+stat_smooth(method = "lm", se = F) + geom_abline(linetype = "dashed")

# It looks like Manganese is the strongest variable and predictor for Chlrophyll a based on the models I ran to determine best fit and minimum variance in residuals.  

```


```{r}
#Multiple Correlation
pairs(~wq2com$`Chlorophyll a`+ wq2com$Depth + wq2com$`Manganese (Dissolved)` + wq2com$`Ammonia (Total)`, data = wq2com)

cov(wq2com[c(8)])
cor(wq2com[c(8)])

```


```{r}

rmse <- function(obs, pred) {sqrt(mean((obs-pred)^2))}
mae <- function(obs, pred) {mean(abs(obs-pred))}
pbias <- function(obs, pred) {100*(sum(pred-obs)/sum(obs))}

man_stats <- t(tibble(rmse = rmse(wq2com$`Manganese (Dissolved)`,exp(wq2com$CvM)), 
                     mae = mae(wq2com$`Manganese (Dissolved)`,exp(wq2com$CvM)), 
                     pbias = pbias(wq2com$`Manganese (Dissolved)`,exp(wq2com$CvM)), 
                     cor = cor(wq2com$`Manganese (Dissolved)`,exp(wq2com$CvM)))

chl_stats <- t(tibble(rmse = rmse(wq2com$`Chlorophyll a`,wq2com$`Chlorophyll a`.fromC), 
                     mae = mae(wq2com$`Chlorophyll a`,wq2com$`Chlorophyll a`.fromC), 
                     pbias = pbias(wq2com$`Chlorophyll a`,wq2com$`Chlorophyll a`.fromC), 
                     cor = cor(wq2com$`Chlorophyll a`,wq2com$`Chlorophyll a`.fromC)))
stats <- cbind(load.stats, conc.stats)

colnames(stats) <- c("Manganese", "Chlorophyll")
stats


```

Part II

Parallel Regression- Parallel regression specifies an additive model, where there is one regression line (for the continuous predictor) which is displaced vertically according to the mean value of the discrete predictor. It hypothesizes that the only effect of the discrete predictor is to adjust the mean, but that the relation between the continuous predictor and the response variable is then the same for all classes of the discrete predictor.

Using your mean monthly water quality dataset for CA WY 2005-2012, create a new discrete value called "season", and assign a label to your data for "wet season" and "dry season" (you will have decide what constitutes wet and dry). Model the mean monthly Chl-a concentration as a function of season and the water quality variable you idenitifed as MOST important in explaining the variance of Chl-a from question 1. 

Model them first separately, then as an additive model, e.g.:
lm(Chl-a ~ season)
lm(Chl-a ~ my_most_important_variable)
lm(Chl-a ~ season + my_most_important_variable)
How much of the variation in Chl-a is explained by season? By your most important variable? By both together? Is the combined model better than individual models? 
In the parallel regression model (season and your most important variable as predictors), what are the differences in the means between seasons? What is the slope of the linear regression after accounting for season? How does this compare with the slope of the linear regression not considering season?
Are the residuals normally distributed? 
Plot Chl-a vs your most important variable, with the observations colored by season. Add the paralell regression lines from the combined model in the appropraite colors, and add the univariate regression line.


library(lubridate)
require(readxl)
require(broom)
require(car)

install.packages("broom")
install.packages("car")

```{r}
#Wet Season November-May Dry Season June-October
season <- ifelse(wq2com$SampleDate >=11|wq2com$SampleDate<=6, "wet", "dry")
as.data.frame(season)
as.character(season)

as.character(wq2com$`Chlorophyll a`)
```

```{r}
#	Model them first separately, then as an additive model 

#lm(Chl-a ~ season)
lm.Seas<-lm(wq2com$wq.season~wq2com$`Chlorophyll a`, na.rm=T)

#lm(Chl-a ~ my_most_important_variable)
lm.Man<-lm(wq2com$`Manganese (Dissolved)`~wq2com$`Chlorophyll a`, data = wq2com)
summary(lm.Man)
#lm(Chl-a ~ season + my_most_important_variable)

lm.Man.Seas <- lm(wq2com$`Chlorophyll a` ~ wq2com$`Manganese (Dissolved)` + wq2com$wq.season, data = wq2com); summary(lm.chla.season.tmp)

```

```{r}

var.man <- var(wq2com$`Chlorophyll a`, wq2com$`Manganese (Dissolved)`, na.rm=T)
var.chla.tmp

var.man.season <- var(wy.select$`Chlorophyll a`, wy.select$season)
var.chla.season

```

```{r}

ggplot(wq2com, aes(x = wq2com$wq.season, y = wq2com$`Chlorophyll a`) + geom_point()
       
```