---
title: "Comments for S. Soltau HW-7 ES207"
author: Cininta Pertiwi
output: html_notebook
---

I reference the codes in your .Rmd file by the line number of the code.

#### **Part I**

##### *Selecting variables*

I like how you started with looking at the variables from the very beginning and narrowing down to a couple based on your knowledge. It may be helpful to filter the dataset from missing values beforehand so that you would not need to create plots and have them come up empty (%1 light depth, pH, and oxygen).

##### *Data wrangling*

* Line 49 and 50 -- It may be less risky to use a conditional code to filter the rows of interest. Using index to select the rows can be hard to do when the dataset is big. Below is an example of using the `filter()` function to select rows that are between October 2004 to July 2012. With `filter()` and using piping, it avoids having the two separate datasets such as the wq1 and wq2 in your code.

```{r eval = FALSE}
# code not run

wq_bymonth <- wq_bymonth %>%
  filter(Year > 2003) %>%                 # select observations after 2003
  filter(!(Year==2004 && Month<10)) %>%   # take out January-September 2004
  filter(!(Year==2012 && Month>9))        # take out October-December 2012
```

* Line 77 -- In your code it did not look like the month and year were correctly separated (in the tibble the Month column had the date wile the Day and Year column were empty. The `lubridate` package can separate out months and years from a date format. By using `lubridate`, the `mutate()` function, and piping the month and year are taken out of the date format and put into their own columns.

```{r eval=FALSE}
# code not run

wq_bymonth <- wq %>%
  mutate(Year = lubridate::year(SampleDate)) %>%    # add Year column
  mutate(Month = lubridate::month(SampleDate))      # add Month column
```

##### *Other*

* I think my comments for this part are only for the data wrangling steps. Otherwise I think your steps for deciding Manganese as the most important variable were clear and well-reasoned although I am curious though on why and how you decided to do a log transform on the data.
* I was not able to get the outputs from the last chunk in this part (lines 174 - 195). The part that was signaling an error were lines 185 - 188 for the variable `chla_stats`. It was because of the `.fromC` after each line. I tried to see why this was but I was not clear on what the code was meant to do.

```{r eval=FALSE}
# code not run

# this was the code that had issues due to the .fromC at the end
chl_stats <- t(tibble(rmse = rmse(wq2com$`Chlorophyll a`,wq2com$`Chlorophyll a`.fromC), 
                     mae = mae(wq2com$`Chlorophyll a`,wq2com$`Chlorophyll a`.fromC), 
                     pbias = pbias(wq2com$`Chlorophyll a`,wq2com$`Chlorophyll a`.fromC), 
                     cor = cor(wq2com$`Chlorophyll a`,wq2com$`Chlorophyll a`.fromC)))
```


#### **Part II**

##### *Data wrangling*

* For Part II, it was asked that the data is classified as wet or dry season. There was a `season` variable created in the code which returned a vector containing 'wet' and 'dry' but it was not joined together with the other variables. It may have been more helpful if instead of storing the wet or dry conditional in a standalone variable, it could have been stored directly as an additional column to the existing tibble with all the other variables.

```{r eval = FALSE}
# code not run

# line 223 in code:
season <- ifelse(wq2com$SampleDate >=11|wq2com$SampleDate<=6, "wet", "dry")

# instead try something such as below so that it becomes another column joined with the other variables
name_of_existing_tibble$season <- ifelse(wq2com$SampleDate >=11|wq2com$SampleDate<=6, "wet", "dry")
```

##### *Other*

* I'm not sure if this was just on my side but I was not able to get outputs from the chunks in Part II. Initially I thought it may have had to do with the wrangling of the season category but then I looked at the code and it may be because it was not able to find some of the variables in the code. The variable `wq.season` was called in lines 234, 241, and 257 but it does seem to have been previously created. the same goes for variable `wy.select` in line 250.
